---
title: "Convert zip codes to FIPS codes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load the required packages.

```{r message = FALSE}
library(tidyverse)
library(here)
library(zipcodeR)
```


Obtain county data from US Census and cache.

```{r message = FALSE}
library(tidycensus)
options(tigris_use_cache = TRUE)

if (FALSE) {
  # download all US counties from US Census
  counties_raw <- get_acs(geography = "county", variables = "B19013_001")
  write_csv(counties_raw, here("data", "counties_census_data.csv"))
}

```


Obtain raw data.

```{r message = FALSE}
# conversion table from state names to state codes
state_lookup <- read_csv(here("data", "state_lookup.csv"))

# read in census data and clean
counties <- read_csv(here("data", "counties_census_data.csv")) %>%
  extract(NAME, into = c("county", "state"), "(.*), (.*)") %>%
  rename(FIPS = GEOID) %>%
  left_join(state_lookup, by = "state") %>%
  mutate(
    # always spell city with capital C
    county = sub("city", "City", county)
  )
```

The following code needs to be in a separate chunk so we can suppress all the output it generates.

```{r eval = FALSE}
# get info on all available zip codes
zip2county_raw <- reverse_zipcode(sprintf("%05i", 1:99999)) %>%
  select(zipcode, county, state_code = state)
```

```{r include = FALSE}
# get info on all available zip codes
zip2county_raw <- reverse_zipcode(sprintf("%05i", 1:99999)) %>%
  select(zipcode, county, state_code = state)
```

The county names between the two data sources differ somewhat. So we need to fix up the zip code lookup table so it matches the naming/spelling in the counties table. Also, there are cases where no county info is available at all, and we can fix that manually as well.

```{r}
zip2county <- zip2county_raw %>%
  mutate(
    # always spell city with capital C
    county = sub("city", "City", county)
  ) %>%
  mutate(
    # manual fix-up of various spelling and naming mistakes
    county = case_when(
      # first we do obvious misspellings
      county == "St Clair County" & state_code == "AL" ~ "St. Clair County",
      county == "Municipality of Anchorage" & state_code == "AK" ~ "Anchorage Municipality",
      county == "City and Borough of Juneau" & state_code == "AK" ~ "Juneau City and Borough",
      county == "Desoto County" & state_code == "FL" ~ "DeSoto County",
      county == "Dekalb County" & state_code == "GA" ~ "DeKalb County",
      county == "Dupage County" & state_code == "IL" ~ "DuPage County",
      county == "Mchenry County" & state_code == "IL" ~ "McHenry County",
      county == "Mclean County" & state_code == "IL" ~ "McLean County",
      county == "La Salle County" & state_code == "IL" ~ "LaSalle County",
      county == "La Porte County" & state_code == "IN" ~ "LaPorte County",
      county == "Lagrange County" & state_code == "IN" ~ "LaGrange County",
      county == "Mccracken County" & state_code == "KY" ~ "McCracken County",
      county == "St Bernard Parish" & state_code == "LA" ~ "St. Bernard Parish",
      county == "St Mary Parish" & state_code == "LA" ~ "St. Mary Parish",
      county == "St Tammany Parish" & state_code == "LA" ~ "St. Tammany Parish",
      county == "St Martin Parish" & state_code == "LA" ~ "St. Martin Parish",
      county == "St Landry Parish" & state_code == "LA" ~ "St. Landry Parish",
      county == "St James Parish" & state_code == "LA" ~ "St. James Parish",
      county == "La Salle Parish" & state_code == "LA" ~ "LaSalle Parish",
      county == "St Mary's County" & state_code == "MD" ~ "St. Mary's County",
      county == "St Louis County" & state_code == "MN" ~ "St. Louis County",
      county == "St Louis County" & state_code == "MO" ~ "St. Louis County",
      county == "St Francois County" & state_code == "MO" ~ "St. Francois County",
      county == "Dekalb County" & state_code == "MO" ~ "DeKalb County",
      county == "Desoto County" & state_code == "MS" ~ "DeSoto County",
      
      county == "Coös County" & state_code == "NH" ~ "Coos County",

      county == "Doa Ana County" & state_code == "NM" ~ "Doña Ana County",
      county == "Dona Ana County" & state_code == "NM" ~ "Doña Ana County",
      county == "Mckinley County" & state_code == "NM" ~ "McKinley County",
      
      county == "St Lawrence County" & state_code == "NY" ~ "St. Lawrence County",
      county == "Mcminn County" & state_code == "TN" ~ "McMinn County",
      
      county == "Mclennan County" & state_code == "TX" ~ "McLennan County",
      county == "De Witt County" & state_code == "TX" ~ "DeWitt County",
      
      county == "James City" & state_code == "VA" ~ "James City County",
      county == "Charles City" & state_code == "VA" ~ "Charles City County",
            
      county == "San Juan County" & state_code == "PR" ~ "San Juan Municipio",
      county == "San Juan" & state_code == "PR" ~ "San Juan Municipio",
      county == "Carolina Municipio" & state_code == "PR" ~ "Carolina Municipio",
      county == "Gunica Municipio" & state_code == "PR" ~ "Guánica Municipio",
      county == "Las Maras Municipio" & state_code == "PR" ~ "Las Marías Municipio",
      county == "Peuelas Municipio" & state_code == "PR" ~ "Peñuelas Municipio",
      county == "Bayamn Municipio" & state_code == "PR" ~ "Bayamón Municipio",
      county == "Bayamón" & state_code == "PR" ~ "Bayamón Municipio",
      county == "Catao Municipio" & state_code == "PR" ~ "Cataño Municipio",
      county == "Cataño" & state_code == "PR" ~ "Cataño Municipio",
      county == "Rincn Municipio" & state_code == "PR" ~ "Rincón Municipio",
      county == "Mayagez Municipio" & state_code == "PR" ~ "Mayagüez Municipio",
      county == "San Germn Municipio" & state_code == "PR" ~ "San Germán Municipio",
      county == "San Sebastin Municipio" & state_code == "PR" ~ "San Sebastián Municipio",
      county == "Canvanas Municipio" & state_code == "PR" ~ "Canóvanas Municipio",
      county == "Ro Grande Municipio" & state_code == "PR" ~ "Río Grande Municipio",

      county == "Manat Municipio" & state_code == "PR" ~ "Manatí Municipio",
      county == "Comero Municipio" & state_code == "PR" ~ "Comerío Municipio",
      county == "Juana Daz Municipio" & state_code == "PR" ~ "Juana Díaz Municipio",

      county == "Aasco Municipio" & state_code == "PR" ~ "Añasco Municipio",
      county == "Loza Municipio" & state_code == "PR" ~ "Loíza Municipio",
      county == "Toa Alta" & state_code == "PR" ~ "Toa Alta Municipio",
      county == "Carolina" & state_code == "PR" ~ "Carolina Municipio",

      # now more complicated cases, regions that were renamed or otherwise changed
            
      county == "Wade Hampton Census Area" & state_code == "AK" ~ "Kusilvak Census Area",

      # the following two are not entirely correct, regions overlap but aren't exactly the same between Petersburg Borough and Prince of Wales-Hyder Census area
      county == "Petersburg Census Area" & state_code == "AK" ~ "Petersburg Borough",
      county == "Prince of Wales-Outer Ketchikan Borough" & state_code == "AK" ~ "Prince of Wales-Hyder Census Area",

      county == "Shannon County" & state_code == "SD" ~ "Oglala Lakota County",

      # everything else carries over unchanged
      TRUE ~ county
    )
  ) %>%
  mutate(
    # fix up a few more cases manually that aren't correct in the raw database
    county = case_when(
      zipcode == "99623" ~ "Matanuska-Susitna Borough",
      zipcode == "98189" ~ "King County",
      zipcode == "98082" ~ "Snohomish County",
      zipcode == "97475" ~ "Lane County",
      zipcode == "87547" ~ "Los Alamos County",
      zipcode == "84129" ~ "Salt Lake County",
      zipcode == "75033" ~ "Collin County", # (mostly, but partially in Denton County, TX)
      #zipcode == "73047" ~ "" # spans multiple counties in OK
      zipcode == "72255" ~ "Pulaski County",
      zipcode == "68544" ~ "Lancaster County",
      zipcode == "64162" ~ "Clay County", # (may also be in Jackson County)
      zipcode == "60569" ~ "DuPage County",
      zipcode == "56999" ~ "District of Columbia",
      zipcode == "50983" ~ "Polk County",
      zipcode == "50982" ~ "Polk County",
      zipcode == "46213" ~ "Marion County",
      zipcode == "46210" ~ "Marion County",
      zipcode == "46197" ~ "Hendricks County",
      zipcode == "41025" ~ "Kenton County",
      zipcode == "41021" ~ "Boone County",
      zipcode == "40166" ~ "Bullitt County",
      zipcode == "35270" ~ "Jefferson County",
      zipcode == "34441" ~ "Citrus County",
      zipcode == "33206" ~ "Miami-Dade County",
      zipcode == "33198" ~ "Miami-Dade County",
      zipcode == "33192" ~ "Miami-Dade County",
      zipcode == "33191" ~ "Miami-Dade County",
      zipcode == "33106" ~ "Miami-Dade County",
      zipcode == "31144" ~ "Cobb County",
      zipcode == "30645" ~ "Oconee County",
      zipcode == "22350" ~ "Alexandria City",
      zipcode == "20417" ~ "District of Columbia",
      TRUE ~ county
    )
  )

```

Now we can join the two tables by county and state and write the final result.

```{r}
zip2fips <- zip2county %>%
  left_join(counties, by = c("county", "state_code")) %>%
  select(zipcode, county, state, state_code, FIPS)
write_csv(zip2fips, here("data", "zip2fips.csv"))
```

## Unresolved cases

The unresolved cases have `NA` in the `FIPS` column.

```{r}
unresolved <- zip2fips %>%
  filter(is.na(FIPS))

# number of unresolved cases
nrow(unresolved)
```

All the remaining cases with county names have county-state mismatch. Many (all?) of these are zip codes that span state boundaries and the county is correct on one side while the state is correct on the other side.

Example: 86044 spans Arizona and Utah, and the part on the Arizona side seems to be in Coconino County (according to Google).

```{r}
# unresolved cases with name
unresolved_named <- filter(unresolved, county != "") %>%
  arrange(county) %>%
  select(zipcode, county, state_code)

unresolved_named
```

The remaining cases without county names require manual look-up or may not exist anymore. I have manually reviewed everything except the ones in PR, and I could not locate the remaining ones outside of PR.

```{r}
# unresolved cases without name
unresolved_unnamed <- filter(unresolved, county == "") %>%
  arrange(desc(zipcode)) %>%
  select(zipcode, county, state_code)

unresolved_unnamed
```
